---
title: "CFSS Homework 04"
author: "Cory J. Cascalheira"
date: "December 27, 2018"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
___
Load dependencies.

```{r message=FALSE}
library(tidyverse) # utility
```

## Unique Values
___
Recall the structure of the `iris` data frame.

```{r}
str(iris)
```

Determine the number of unique values within each column with a `for` loop. First, solve the problem for one column.

```{r}
length(unique(iris[[1]]))
```

Apply abstraction.

```{r}
# Preallocate space for efficiency
output <- vector(mode = "numeric", length = ncol(iris))

# Iterate over each variable
for (i in seq_along(iris)) {
  output[[i]] <- length(unique(iris[[i]]))
}

# Print number of unique values
output

# Assign names to the output
names(output) <- names(iris)
output
```

Perform the same operation with `map_int()` and an anonymous function.

```{r}
map_int(iris, function(df) length(unique(df)))
```

## Square of Each Element
___
Initialize the vector `x`.

```{r}
x <- seq(from = 1, to = 30)
x
```

Solve the problem for the third value in the vector `x`.

```{r}
x[[3]]^2
```

Apply abstraction.

```{r}
# Preallocate memory for efficiency
output <- vector(mode = "numeric", length = length(x))

# Iterate over each element
for (i in seq_along(x)) {
  output[[i]] <- x[[i]]^2
}

# Print output
output
```

Perform the same operation using `map_int()`.

```{r}

```










# Session Info
___
```{r}
devtools::session_info()
```