---
title: "CFSS Homework 04"
author: "Cory J. Cascalheira"
date: "December 27, 2018"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/GitHub/cfss-hw04/")
```

# Part 1
___
Load dependencies.

```{r message=FALSE}
library(tidyverse) # utility
```

For the second part of this assignment, I will need to use relative paths. If I had created an R project, this would have been simple since the working directory would have been the project folder. I did not, so let's set the working directory now. In an R Markdown document, `setwd()` does not work. Instead, we must set the knitr option in the setup chunk using `knitr::opts_knit$set(root.dir = "~/GitHub/cfss-hw04/")`.

## Unique Values
___
Recall the structure of the `iris` data frame.

```{r}
str(iris)
```

Determine the number of unique values within each column with a `for` loop. First, solve the problem for one column.

```{r}
length(unique(iris[[1]]))
```

Apply abstraction.

```{r}
# Preallocate space for efficiency
output <- vector(mode = "numeric", length = ncol(iris))

# Iterate over each variable
for (i in seq_along(iris)) {
  output[[i]] <- length(unique(iris[[i]]))
}

# Print number of unique values
output

# Assign names to the output
names(output) <- names(iris)
output
```

Perform the same operation with `map_int()` and an anonymous function.

```{r}
map_int(iris, function(df) length(unique(df)))
```

## Square of Each Element
___
Initialize the vector `x`.

```{r}
x <- seq(from = 1, to = 30)
x
```

Solve the problem for the third value in the vector `x`.

```{r}
x[[3]]^2
```

Apply abstraction.

```{r}
# Preallocate memory for efficiency
output <- vector(mode = "numeric", length = length(x))

# Iterate over each element
for (i in seq_along(x)) {
  output[[i]] <- x[[i]]^2
}

# Print output
output
```

Perform the same operation using `map_dbl()` and an anonymous function.

```{r}
map_dbl(x, function(x) x^2)
```

## Pythagorean Theorem
___
Given the sides of *a* and *b*, find the hypotenuse *c*.

This problem can be solved in cumulative steps. 

The hypotenuse is equal to:

$$c = \sqrt{a^2+b^2}$$

We can use a simple 3-4-5 triangle to make sure our code works. Let's say we know that *a* = 3 and *b* = 4. First, solve the problem with real numbers.

```{r}
sqrt(
  3^2 + 4^2
)
```

Next, plug in variables.

```{r}
# Sides
a = 3
b = 4

# Solve for hypotenuse
c <- sqrt(
  a^2 + b^2
) 

# Print the hypotenuse
c
```

### Simple Function

Here is a simple function that works when we supply both *a* and *b*.

```{r}
# Create function
pythagorean <- function (a, b) {
  c <- sqrt(
    a^2 + b^2
  )
  c
}

# Test function
pythagorean(a = a, b = b)
```

### Complex Function

If we know the hypotenuse and either *a* or *b*, the equation changes:

$$a = \sqrt{c^2 - b^2} \\ b = \sqrt{c^2 - a^2}$$

Let's incorporate this conditional difference into our function. Instead of setting a, b, or c to `NULL`, we can use `missing()` inside of an `if()` statement to [specify optional parameters](https://stackoverflow.com/questions/28370249/correct-way-to-specifiy-optional-arguments-in-r-functions).

```{r}
# Create complex function
pythagorean <- function (a, b, c) {
  
  # Solve for a
  if (missing(a)) {
    a <- sqrt(
      c^2 - b^2
    )
    a
  
  # Solve for b
  } else if (missing(b)) {
    b <- sqrt(
      c^2 - a^2
    )
    b
  
  # Solve for c
  } else {
    c <- sqrt(
      a^2 + b^2
    )
    c
  }
}

# Test function on random values
pythagorean(a = 3, b = 4)
pythagorean(a = 6, c = 10)
pythagorean(b = 7, c = 11)
```

### Check for Errors

Now let's add some error checking to make our function more robust. If we want this R markdown document to [knit properly with errors in the chunks](https://stackoverflow.com/questions/34137797/rmarkdown-run-code-and-display-errors-in-document), we must set the option `error=TRUE`.

Without making any changes, we see that our function will throw an error if only one argument is supplied.

```{r error=TRUE}
pythagorean(a = 2)
```

This is really convenient.

```{r error=TRUE}
pythagorean(b = 2)
```

R is smart enough to know that the equations cannot be evaluated with a missing variable.

```{r error=TRUE}
pythagorean(c = 2)
```

Thus, all we must do is specify a `stop()` condition for when a user: 

* supplies all three arguments at once; or
* supplies a non-numeric argument.

```{r}
# Create complex function
pythagorean <- function (a, b, c) {
  
  # Check for all three arguments at once
  if (!is.na(a) && !is.na(b) && !is.na(c)) {
    stop("Cannot supply all sides of the triangle.")
  }
  
  # Check for non-numeric arguments
  if (!is.numeric(a) || !is.numeric(b) || !is.numeric(c)) {
    stop("Arguments must be numerics.")
  }
  
  # Solve for a
  if (missing(a)) {
    a <- sqrt(
      c^2 - b^2
    )
    a
  
  # Solve for b
  } else if (missing(b)) {
    b <- sqrt(
      c^2 - a^2
    )
    b
  
  # Solve for c
  } else {
    c <- sqrt(
      a^2 + b^2
    )
    c
  }
}
```

What happens if we supply all three arguments at once?

```{r error=TRUE}
pythagorean(a = 2, b = 5, c = 7)
```

What if we supply the English equivalent of 2?

```{r error=TRUE}
pythagorean(a = "two", b = 5)
```

Our function is not evaluating the non-numeric `if()` statement. Will this behavior change if we set all values to `NULL`?

This is a workaround that requires `is.null()` instead of `missing()`. Without `missing()`, the addition of an error check for when a user supplies only one side of the triangle is necessary.

The error check for all arguments at once will stay the same.

In order to properly evaluate whether an argument is non-numeric, the argument cannot be missing (i.e., cannot be set to `NULL`).

```{r}
# Create complex function
pythagorean <- function (a = NULL, b = NULL, c = NULL) {
  
  ### Error Checking ###
  
  # Check for single arguments
  if ((is.null(a) && is.null(b)) || (is.null(b) && is.null(c)) || (is.null(a) && is.null(c))) {
    stop("Must supply at least two arguments.")
  }
  
  # Check for all three arguments at once
  if (!is.null(a) && !is.null(b) && !is.null(c)) {
    stop("Cannot supply all sides of the triangle.")
  }
  
  # Check for non-numeric argument in a
  if (!is.numeric(a) && !is.null(a)) {
    stop("Argument `a` must be numeric.")
  }
  
  # Check for non-numeric argument in b
  if (!is.numeric(b) && !is.null(b)) {
    stop("Argument `b` must be numeric.")
  }
  
  # Check for non-numeric argument in c
  if (!is.numeric(c) && !is.null(c)) {
    stop("Argument `c` must be numeric.")
  }
  
  ### Pythagorean Theorem ###
  
  # Solve for a
  if (is.null(a)) {
    a <- sqrt(
      c^2 - b^2
    )
    a
    
    # Solve for b
  } else if (is.null(b)) {
    b <- sqrt(
      c^2 - a^2
    )
    b
    
    # Solve for c
  } else {
    c <- sqrt(
      a^2 + b^2
    )
    c
  }
}
```

Does `pythagorean()` throw an error when the only one side of the triangle is specified?

```{r error=TRUE}
pythagorean(a = 3)
```

Does `pythagorean()` throw an error if all three arguments are specified?

```{r error=TRUE}
pythagorean(2, 6, 11)
```

Does `pythagorean()` throw an error if the argument is non-numeric?

```{r error=TRUE}
pythagorean(b = 3, c = "eleven")
```

# Part 2
___
I downloaded 9 of the economic indicators directly from the [World Bank](https://data.worldbank.org/indicator). After unzipping the files, I placed each `csv` into my `data_world_bank` folder.

First, let's find all the file names in the appropriate folder. We don't care about the names *per se* since they all end in `csv`. In the call to `dir()`, remember to escape the period.

```{r}
files <- dir(path = "data_world_bank/", pattern = "\\.csv$", full.names = TRUE)
files
```

## Create Function
___
Each file has blank space at the top, so before we run an iterative function on each file, let's practice importing the information for one file.

```{r message=FALSE, warning=FALSE}
dummy <- map("data_world_bank/API_BM.GSR.ROYL.CD_DS2_en_csv_v2_10232025.csv", read_csv, skip = 2)
head(dummy[[1]])
```

Our function can rename the variables and tranforms each data frame into long format.

```{r}
dummy[[1]] %>%
  # Remove empty variables
  select(-starts_with("X")) %>%
  # Rename columns using snake_case
  rename(
    "country_name" = "Country Name",
    "country_code" = "Country Code",
    "indicator_name" = "Indicator Name",
    "indicator_code" = "Indicator Code"
  ) %>%
  # Gather years into single column
  gather(key = year, value = amount, -c(1:4))
```

At this point, we are ready to apply abstraction to the steps so far.

```{r}
condense <- function (files) {
  
  # Read all csv into a list
  file_list <- files %>%
    map(read_csv, skip = 2)
  
  # Prepare new list
  output <- vector(mode = "list", length = length(files))
  
  # Clean each data frame
  for (i in seq_along(files)) {
    output[[i]] <- file_list[[i]] %>%
      
      # Remove empty variables
      select(-starts_with("X")) %>%
      
      # Rename columns using snake_case
      rename(
        "country_name" = "Country Name",
        "country_code" = "Country Code",
        "indicator_name" = "Indicator Name",
        "indicator_code" = "Indicator Code"
      ) %>%
      
      # Gather years into single column
      gather(key = year, value = amount, -c(1:4))
  }
  
  output
  
}
```

Does the function work as expected?

```{r message=FALSE, warning=FALSE}
(trial <- condense(files = files))
```

Now we want the function to combine each nested data frame into one master tibble. [We could](https://stackoverflow.com/questions/2851327/convert-a-list-of-data-frames-into-one-data-frame) combine a `do.call()` with `rbind()`. There are also other, deprecated options such as `plyr::rbind.fill` or `rbind_all()`.

Instead, we can simply call `bind_rows()`.

```{r}
bind_rows(trial)
```

Let's incorporate this small change into our function.

```{r}
condense <- function (files) {
  
  # Read all csv into a list
  file_list <- files %>%
    map(read_csv, skip = 2)
  
  # Prepare new list
  output <- vector(mode = "list", length = length(files))
  
  # Clean each data frame
  for (i in seq_along(files)) {
    output[[i]] <- file_list[[i]] %>%
      
      # Remove empty variables
      select(-starts_with("X")) %>%
      
      # Rename columns using snake_case
      rename(
        "country_name" = "Country Name",
        "country_code" = "Country Code",
        "indicator_name" = "Indicator Name",
        "indicator_code" = "Indicator Code"
      ) %>%
      
      # Gather years into single column
      gather(key = year, value = amount, -c(1:4))
  }
  
  # Bind all data frames into master tibble
  bind_rows(output)
  
}
```

And try it.

```{r warning=FALSE, message=FALSE}
(indicators <- condense(files))
```

## Brief Report
___
Since we will produce graphs, let's set the default theme.

```{r}
theme_set(theme_minimal())
```

Review indicators to analyze.

```{r}
unique(indicators$indicator_name)
```

Let's analyze the central government debt and GDP.

```{r}
debt_gdp <- indicators %>%
  # Select indicators of interest
  filter(indicator_name == "Central government debt, total (% of GDP)" |
         indicator_name == "GDP (current US$)") %>%
  # Remove missing values
  filter(!is.na(amount))
```

We can simplify the name of each indicator and transform `year` into an integer.

```{r}
# Simplify indicator
debt_gdp$indicator_name <- fct_recode(factor(debt_gdp$indicator_name),
  "debt" = "Central government debt, total (% of GDP)",
  "gdp" = "GDP (current US$)")

# Year as integer
debt_gdp$year <- as.integer(debt_gdp$year)
```

### Exploratory Analysis

What does GDP look like for the world?

```{r}
debt_gdp %>%
  filter(country_name == "World") %>%
  # Scale large number
  mutate(amount = amount / 1000000000000) %>%
  ggplot(aes(x = year, y = amount)) +
    geom_line(color = "cadetblue", size = 1) +
    # Set custom breaks
    scale_x_continuous(breaks = seq(1960, 2017, by = 5), expand = c(0, 0)) +
    # Add dollar sign to values along y-axis
    scale_y_continuous(labels = scales::dollar) +
    labs(title = "Global Gross Domestic Product (GDP)",
         subtitle = "1960 - 2017",
         caption = "Source: World Bank",
         y = "Amount (Trillions)",
         x = "Year")
```

What does GDP look like for the United States, China, and the European Union?

```{r}
# Vector of country names
countries <- c("United States", "China", "European Union")

# Visualize GDP 
debt_gdp %>%
  filter(country_name %in% countries & indicator_name == "gdp") %>%
  # Scale large number
  mutate(amount = amount / 1000000000000) %>%
  ggplot(aes(x = year, y = amount, color = country_name)) +
    geom_line(size = 1) +
    scale_color_brewer(palette = "Paired") +
    # Set custom breaks
    scale_x_continuous(breaks = seq(1960, 2017, by = 5), expand = c(0, 0)) +
    guides(color = guide_legend(title = "Country")) +
    # Add dollar sign to values along y-axis
    scale_y_continuous(labels = scales::dollar) +
    labs(title = "Gross Domestic Product (GDP)",
         subtitle = "1960 - 2017",
         caption = "Source: World Bank",
         y = "Amount (Trillions)",
         x = "Year") +
    # Move legend to top
    theme(legend.position = "top")
```

### U.S. vs U.K.

Let's compare the GDP between the United States and United Kingdom.

```{r}
# Vector of country names
western <- c("United States", "United Kingdom")

# Visualize GDP 
debt_gdp %>%
  filter(country_name %in% western & indicator_name == "gdp") %>%
  # Scale large number
  mutate(amount = amount / 1000000000000) %>%
  ggplot(aes(x = year, y = amount, color = country_name)) +
    geom_line(size = 1) +
    scale_color_brewer(palette = "Paired") +
    # Set custom breaks
    scale_x_continuous(breaks = seq(1960, 2017, by = 5), expand = c(0, 0)) +
    guides(color = guide_legend(title = "Country")) +
    # Add dollar sign to values along y-axis
    scale_y_continuous(labels = scales::dollar) +
    labs(title = "Gross Domestic Product (GDP)",
         subtitle = "1960 - 2017",
         caption = "Source: World Bank",
         y = "Amount (Trillions)",
         x = "Year") +
    # Move legend to top
    theme(legend.position = "top")
```

How do the United States and United Kingdom compare on debt? Remember, the debt is measured as percent of GDP.

```{r}
# Visualize GDP 
debt_gdp %>%
  filter(country_name %in% western & indicator_name == "debt") %>%
  mutate(amount = amount / 100) %>%
  ggplot(aes(x = year, y = amount, color = country_name)) +
    geom_line(size = 1) +
    scale_y_continuous(limits = c(0, 1.3), labels = scales::percent, expand = c(0, 0)) +
    # Set custom breaks
    scale_x_continuous(breaks = seq(1970, 2018, by = 3), expand = c(0, 0)) +
    scale_color_brewer(palette = "Paired") +
    guides(color = guide_legend(title = "Country")) +
    labs(title = "National Debt",
         caption = "Source: World Bank",
         y = "Percent of GDP",
         x = "Year") +
    # Move legend to top
    theme(legend.position = "top")
```

The debt for the United States has reached 100% of GDP by 2017, but the nation's GDP continues to climb in what appears to be an exponential trend. The United Kingdom, on the other hand, reached a debt that was 120% of GDP in 2017. Growth appears to have plateaued. The graph on national debt evinces a sharp rise in debt for both countries after the global recession of 2008.

# Session Info
___
```{r}
devtools::session_info()
```